\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-10-06 adaptive-estimation.tex
corrected Isoprimetric Lemma. Went through it with Vardan

We have the following results:
\begin{thm} \label{thm:sgd}
Consider the sequence $\left\{\bar{\theta}_n \right\}_{n\in \mathbb N}$ defined by \eqref{eq:sgd_est}. 
\begin{enumerate}
\item[(i)] Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
\begin{equation} \label{eq:conditions1}
\begin{cases}
\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n), &  \\
\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty, & 
\mathrm{for~some~}0< \lambda \leq 1
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). Then
\[
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left(0,\frac{1}{\eta(0)}\right).
\]
%
%
\item [(ii)] Assume in addition that $f(x)$ is continuously differentiable with a finite location Fisher information, namely
\[
I_\theta = \int_{\mathbb R} \left( \frac{f'(x)}{f(x)}  \right)^2 f(x) dx < \infty.
\]
%(or that $f(x-\theta)$ is differentiable in quadratic mean).  
Then for any bounded, symmetric, and quasi-convex function $L$, 
\begin{align} 
& \liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  \mathbb E \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
\\
& \qquad \qquad \qquad \qquad = \mathbb E \left[ L (Z/\sqrt{\eta(0)}) \right], \label{eq:attaining_LAM}
\end{align}
where $Z \sim \Ncal(0,1)$. 
%
%
\item[(iii)] Assume that, in addition to \eqref{eq:conditions1}, $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
\begin{equation} \label{eq:conditions2}
\begin{cases}
\gamma_n = o(n^{-2/3}), &  \\
\sum_{n=1}^\infty \gamma_n = \infty.  & \\
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ with $\beta \in (2/3,1)$). Then
\[
 \ex{\left( \bar{\theta}_n - \theta \right)^2} = \frac{1}{n\eta(0)} + o(n^{-1}). 
\]
\end{enumerate}
\end{thm}

@@

\subsection{Maximal Efficiency using One Round of Threshold Adaptation}
In Section~\ref{sec:preliminary} we considered an estimator that is based on binary messages of the form 
\[
B_i = \mathbf 1_{X_i > \theta_0},\qquad i=1,\ldots,n,
\]
and deduced that it is asymptotically normal with variance $1/\eta(\theta-\theta_0)$.
%
We now show that a similar encoding leads to an asymptotically normal estimator attaining the lower variance bound $1/\eta(0)$, provided we allow to  update the threshold value $\theta_0$ based on previously observed bits at least once. 
%
In this procedure we separate the sample into two disjoint sets: $X_1,\ldots,X_{n_1}$ and $X_{n_1+1},\ldots,X_n$ for some $n_1 < n$.
We first use the estimator \eqref{eq:estimator_naive} to obtain an estimate ${\theta}_{n_1}$ based on $B_1,\ldots,B_{n_1}$, and then use ${\theta}_{n_1}$ as the new threshold value to obtain messages $B_{n_1+1}, \ldots, B_n$. Figure~\ref{fig:one_round} illustrates a diagram of this procedure. 
%
The specific encoding and estimation scheme, as well as its asymptotic performance, are given by the following theorem:
\begin{thm}
For $i=1,\ldots,n$ set
\[
B_i = \begin{cases}
 \mathbf 1_{X_i \geq \theta_0} & i = 1,\ldots,n_1, \\
 \mathbf 1_{X_i \geq {\theta}_{n_1} }& i={n_1+1,\ldots,n},
\end{cases}
\]
where $n = n_1+ n_2$ and 
\[
{\theta}_{n_1} \triangleq \theta_0 + F^{-1}\left(
\frac{1}{n_1} \sum_{i=1}^{n_1} B_i 
 \right).
\] 
Let 
\[
{\theta}_{n_2}  = {\theta}_{n_1} +  F^{-1} \left( \frac{1}{n_2} \sum_{i=n_1+1}^{n_2} B_i \right)
\]
and assume that $n_1(n) \rightarrow \infty$ and $n_1(n)/n \rightarrow 0$. Then:
\begin{align*}
 \sqrt{n} \left( {\theta}_{n_2} - \theta  \right) \overset{d}{\longrightarrow}  \Ncal\left( 0, 1/\eta(0) \right).
\end{align*}
\end{thm}
%
%
\begin{proof}
For $t\in \mathbb R$, set
\[
p_{n_2}(t) \triangleq \frac{1}{n_2} \sum_{i=n_1+1}^{n} \mathbf 1_{X_i \geq t}. 
\]
From the central limit theorem
\begin{align*}
& \sqrt{n} \left( p_{n_2}(t) - F( \theta - t) \right)  = \sqrt{\frac{n}{n_2}} \sqrt{n_2} \left(  p_{n_2}(t) - F(\theta - t) \right) \\
& \overset{d}{\rightarrow} \Ncal\left( 0, V(t) \right),
\end{align*}
where
\[
V(t) \triangleq F(\theta - t) F \left( t - \theta\right). 
\]
Applying the delta method to $p_{n_2}(t)$ with $g(x) = F^{-1}(x)$ implies
\begin{align*}
& \sqrt{n} \left( t + F^{-1}(\hat{p}_{n_2}(t)) - \theta \right) \\
&= \sqrt{n} \left( g(\hat{p}_{n_2}(t)) - g \left(F(\theta-t) \right)  \right) \\
& \overset{d}{\to} \Ncal\left(0, 
1/\eta(\theta-t) \right).
\end{align*}
%where this convergence is uniform in $t$ since all moments of $F^{-1}(\hat{p}_{n_2})$ exists. 
On the other hand, by the law of large numbers,
\[
p_{n_1} \triangleq \frac{1}{n_1} \sum_{i=1}^{n_1} B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),
\]
so that ${\theta}_{n_1}$ converges almost surely to $\theta$ as $n$ goes to infinity and thus 
\[
\eta( {\theta}_{n_1}-\theta) \overset{a.s.}{\rightarrow} \eta(0). 
\]
Finally,
\begin{align*}
& \sqrt{n}\left({\theta}_{n_2} - \theta\right) \\
& = \sqrt{n} \left( {\theta}_{n_1} + F^{-1}(\hat{p}_{n_2}({\theta}_{n_1})) - \theta \right)  \\
& \overset{d}{\to} \Ncal\left( 0, 
1/\eta(0) \right) 
\end{align*}
by Slutsky's theorem.
\end{proof}
%

>>>>

We have the following results:
\begin{thm} \label{thm:sgd}
Consider the sequence $\left\{\bar{\theta}_n \right\}_{n\in \mathbb N}$ defined by \eqref{eq:sgd_est}. 
\begin{enumerate}
\item[(i)] Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
\begin{equation} \label{eq:conditions1}
\begin{cases}
\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n), &  \\
\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty, & 
\mathrm{for~some~}0< \lambda \leq 1
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). Then
\[
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left(0,\frac{1}{\eta(0)}\right).
\]
%
%
\item [(ii)] Assume in addition that $f(x)$ is continuously differentiable with a finite Fisher information 
\[
\sigma_f^2 \triangleq \ex{\frac{f'(X)}{f(X)} } 
\]
for the location in the family $\{P_\theta\}$.
%(or that $f(x-\theta)$ is differentiable in quadratic mean).  
Then for any bounded, symmetric, and quasi-convex function $L$, 
\begin{align} 
& \liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  \mathbb E \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
\\
& \qquad \qquad \qquad \qquad = \mathbb E \left[ L (Z/\sqrt{\eta(0)}) \right], \label{eq:attaining_LAM}
\end{align}
where $Z \sim \Ncal(0,1)$. 
%
%
\item[(iii)] Assume that, in addition to \eqref{eq:conditions1}, $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
\begin{equation} \label{eq:conditions2}
\begin{cases}
\gamma_n = o(n^{-2/3}), &  \\
\sum_{n=1}^\infty \gamma_n = \infty.  & \\
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ with $\beta \in (2/3,1)$). Then
\[
 \ex{\left( \bar{\theta}_n - \theta \right)^2} = \frac{1}{n\eta(0)} + o(n^{-1}). 
\]
\end{enumerate}
\end{thm}

@@

\subsection{Maximal Efficiency using One Round of Threshold Adaptation}
In Section~\ref{sec:preliminary} we considered an estimator that is based on binary messages of the form 
\[
B_i = \mathbf 1_{X_i > \theta_0},\qquad i=1,\ldots,n,
\]
and deduced that it is asymptotically normal with variance $1/\eta(\theta-\theta_0)$.
%
We now show that a similar encoding leads to an asymptotically normal estimator attaining the lower variance bound $1/\eta(0)$, provided we allow to  update the threshold value $\theta_0$ based on previously observed bits at least once. 
%
In this procedure we separate the sample into two disjoint sets: $X_1,\ldots,X_{n_1}$ and $X_{n_1+1},\ldots,X_n$ for some $n_1 < n$.
We first use the estimator \eqref{eq:estimator_naive} to obtain an estimate ${\theta}_{n_1}$ based on $B_1,\ldots,B_{n_1}$, and then use ${\theta}_{n_1}$ as the new threshold value to obtain messages $B_{n_1+1}, \ldots, B_n$. Figure~\ref{fig:one_round} illustrates a diagram of this procedure. 
%
The specific encoding and estimation scheme, as well as its asymptotic performance, are given by the following theorem:
\begin{thm}
For $i=1,\ldots,n$ set
\[
B_i = \begin{cases}
 \mathbf 1_{X_i \geq \theta_0} & i = 1,\ldots,n_1, \\
 \mathbf 1_{X_i \geq {\theta}_{n_1} }& i={n_1+1,\ldots,n},
\end{cases}
\]
where $n = n_1+ n_2$ and 
\[
{\theta}_{n_1} \triangleq \theta_0 + F^{-1}\left(
\frac{1}{n_1} \sum_{i=1}^{n_1} B_i 
 \right).
\] 
Let 
\[
{\theta}_{n_2}  = {\theta}_{n_1} +  F^{-1} \left( \frac{1}{n_2} \sum_{i=n_1+1}^{n_2} B_i \right)
\]
and assume that, as $n\to \infty$, $n_1(n) \rightarrow \infty$ and $n_2(n)/n \rightarrow 1$. Then:
\begin{align*}
 \sqrt{n} \left( {\theta}_{n_2} - \theta  \right) \overset{d}{\longrightarrow}  \Ncal\left( 0, 1/\eta(0) \right).
\end{align*}
\end{thm}
%
%
\begin{proof}
By the law of large numbers,
\[
p_{n_1} \triangleq \frac{1}{n_1} \sum_{i=1}^{n_1} B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),
\]
so that ${\theta}_{n_1}$ converges almost surely to $\theta$ as $n$ goes to infinity. 
For $t\in \mathbb R$, set
\[
p_{n_2}(t) \triangleq \frac{1}{n_2} \sum_{i=n_1+1}^{n} \mathbf 1_{X_i \geq t}. 
\]
Since $n_2(n)/n \to 1$, the central limit theorem and Slutsky's theorem imply
\begin{align*}
\sqrt{n} \left( p_{n_2}(t) - F( \theta - t) \right)   \overset{d}{\rightarrow} \sqrt{F(\theta - t) F \left( t - \theta\right)} Z,
\end{align*}
where $Z \sim \Ncal(0,1)$. In addition, Slutsky's theorem imply
\[
\sqrt{F(\theta-\theta_{n_1}) F(\theta_{n_1} - \theta)} Z_1 \overset{d}{\rightarrow}  F(0) Z,  
\]
hence 
\[
\sqrt{n} \left( p_{n_2}( \theta_{n_1}) - F( \theta - \theta_{n_1}) \right)  \overset{d}{\rightarrow} F(0) Z.
\]
Finally, we use the delta method with $g(x) = F^{-1}(x)$ to conclude
\begin{align*}
& \sqrt{n}\left( \theta_{n_2} - \theta \right) = \sqrt{n} \left( \theta_{n_1} + F^{-1}(p_{n_2}(\theta_{n_1})) - \theta \right) \\
&= \sqrt{n} \left( g(p_{n_2}(\theta_{n_1})) - g \left(F(\theta-\theta_{n_1}) \right)  \right) \overset{d}{\to} \Ncal\left(0, 
F^2(0)/ f^2(0) \right).
\end{align*}
\end{proof}
%
